---
title: "Arithmetic"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Arithmetic}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dbMatrix)
```


### Get test data
The test file is a `dgCMatrix`or compressed sparse column matrix representing a single cell gene expression matrix. The file is in the `data` directory of the package.

Let's load the .rds file and preview the object.

```{r}
dgc <- readRDS("../data/dgc.rds")

dplyr::glimpse(dgc)
```

The file contains 634 rows and 624 columns. The rows represent gene names and the columns represent cell names. The values are integers and represent the number of times a gene is detected in a cell. Like most single-cell RNA-seq data, the matrix is sparse.


### Create a dbMatrix object
Let's create a `dbSparseMatrix` object from the above `dgc` object.
```{r}
# Note: by default the constructor creates a dbMatrix object in-memory
dbsm <- createDBMatrix(value = dgc, 
                       db_path = ":temp:", 
                       name = 'visium', 
                       class = "dbSparseMatrix",
                       overwrite = TRUE)

# preview the object
dbsm
```

### Scalar Arithmetic
Performing scalar arithmetic on a `dbMatrix` object is similar to performing scalar arithmetic on a `dgCMatrix` object. 

Note: scalar arithmetic on a `dbSparseMatrix` results in a `dbDenseMatrix`.
```{r} 
dbsm + 1

dbsm * 100
```

### Matrix Arithmetic
`dbMatrix` also supports matrix arithmetic for `dbMatrix` objects that are [conformable](https://en.wikipedia.org/wiki/Conformable_matrix).
```{r}
dbsm + dbsm
```

### Matrix Multiplication

#### Hadamard product
```{r}
dbsm * dbsm
```

#### Matrix product
TODO

### Shutdown
```{r}
dbMatrix::dbDisconnect(dbsm, shutdown = TRUE)
```

### Session Info
```{r}
sessionInfo()
```


