---
title: "Operations"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
number-sections: true
vignette: >
  %\VignetteIndexEntry{Operations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
options(max.print = 10) # Set the maximum number of rows to display
options(width = 80)     # Set the maximum number of columns to display
```

## Introduction

This vignette introduces the generic functions and operations supported by `dbMatrix`. 

## Loading library

```{r, message=F}
library(dbMatrix)
library(Matrix)
```

## dbMatrix generics

`dbMatrix` subclasses currently support basic generic functions listed below with support for more operations coming soon.

|          | dbSparseMatrix | dbDenseMatrix |
|----------|----------------|---------------|
| colSums  | ✅             | ✅            |
| rowSums  | ✅             | ✅            |
| colMeans | ✅             | ✅            |
| rowMeans | ✅             | ✅            |
| colSds   | 🟧             | ✅            |
| rowSds   | 🟧             | ✅            |
| t        | ✅             | ✅            |
| mean     | ✅             | ✅            |
| nrow     | ✅             | ✅            |
| ncol     | ✅             | ✅            |
| dims     | ✅             | ✅            |
| head     | ✅             | ✅            |
| tail     | ✅             | ✅            |
| ...      |                |               |

## dbSparse Matrix Operations

### Get test data
The test file is a `dgCMatrix`or compressed sparse column matrix representing a single cell gene expression matrix. The file is in the `data` directory of the package.

Let's load the .rds file and preview the object.

```{r}
dgc <- readRDS("../data/dgc.rds")

dplyr::glimpse(dgc)
```

```{r}
# create dbSparseMatrix from the same dgc
sparse = createDBMatrix(value = dgc, 
                        db_path = ":temp:", 
                        name = "sparse", 
                        class = "dbSparseMatrix",
                        overwrite = TRUE)


# preview 
# show function aims to emulate the show method for dgCMatrix
head(sparse)
```

### transpose
```{r}
t(sparse)
```

### colMeans
```{r}
dbMatrix::colMeans(sparse)
```


### colSums
```{r}
dbMatrix::colSums(sparse)
```

### rowMeans
```{r}
dbMatrix::rowMeans(sparse)
```

### rowSums
```{r}
dbMatrix::rowSums(sparse)
```

### dim
```{r}
dim(sparse)

dim(dgc)
```

### Check results are equivalent
<details>
  <summary>Click to expand </summary>
  
  ```{r}
  all.equal(dbMatrix::colMeans(sparse), Matrix::colMeans(dgc))
  
  all.equal(dbMatrix::colSums(sparse), Matrix::colSums(dgc))
  
  all.equal(dbMatrix::rowMeans(sparse), Matrix::rowMeans(dgc))
  
  all.equal(dbMatrix::rowSums(sparse), Matrix::rowSums(dgc))

  ```
</details>


## dbDenseMatrix Operations

```{r, eval=TRUE, message=FALSE, warning=FALSE}
dense = sim_dbDenseMatrix()

# preview
dense
```

### transpose
```{r}
t(dense)
```

### colMeans
```{r}
dbMatrix::colMeans(dense)
```

### colSums
```{r}
dbMatrix::colSums(dense)
```

### rowMeans
```{r}
dbMatrix::rowMeans(dense)
```

### rowSums
```{r}
dbMatrix::rowSums(dense)
```

### mean
```{r}
dbMatrix::mean(dense)
```

### dim
```{r}
dim(dense)
```

## Shutdown
```{r}
dbMatrix::dbDisconnect(dense, shutdown = TRUE)
```

## Session Info
```{r eval=TRUE, message=FALSE}
sessionInfo()
```
