---
title: "Operations"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
number-sections: true
vignette: >
  %\VignetteIndexEntry{Operations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

This vignette introduces the different classes of `dbMatrix` and how to work 
with them at a basic level using generic functions.

## Loading library

```{r, message=F}
library(dbMatrix)
library(Matrix)
```

## Working with large matrices with dbMatrix

`dbMatrix` consists of two dbMatrix subclasses that are designed to work with
dense and sparse matrices. Support for compressed sparse column matrices through the `dgCMatrix` class from the `Matrix` package is currently implemented. For more information about sparse matrices, see [the wiki article here ](https://en.wikipedia.org/wiki/Sparse_matrix). 

`dbMatrix` subclasses currently support basic generic operations listed below with support for more operations coming soon.

|          | dbSparseMatrix | dbDenseMatrix |
|----------|----------------|---------------|
| colSums  | ✅             | ✅            |
| rowSums  | ✅             | ✅            |
| colMeans | ✅             | ✅            |
| rowMeans | ✅             | ✅            |
| colSds   | 🟧             | ✅            |
| rowSds   | 🟧             | ✅            |
| t        | ✅             | ✅            |
| mean     | ✅             | ✅            |
| nrow     | ✅             | ✅            |
| ncol     | ✅             | ✅            |
| dims     | ✅             | ✅            |
| head     | ✅             | ✅            |
| tail     | ✅             | ✅            |
| math     | ✅             | ✅            |
| ...      |                |               |

## dbSparse Matrix Operations

We represent sparse matrices using a triplicate format with `i`, `j`,
and `x` numerical vectors. `i` and `j` refer to the row and column indices,
respectively, with specific `x` values or counts.

The `dbSparseMatrix` class stores this ijx matrix (now technicality a
dataframe) with their associated dimension names and dimensions if available. It
does not contain zeros. Zero values are inferred based on the stored
dimensions.

```{r eval=TRUE, message=FALSE, warning=FALSE}
# simulate a dgC matrix
dgc = dbMatrix:::sim_dgc()

# create dbSparseMatrix from the same dgc in memory
sparse = createDBMatrix(value = dgc, name = "sparse", class = "dbSparseMatrix", overwrite = TRUE)

# preview 
# show function aims to emulate the show method for dgCMatrix
sparse
```

### colMeans
```{r}
dbMatrix::colMeans(sparse)

all.equal(unname(dbMatrix::colMeans(sparse)), Matrix::colMeans(dgc))
```

### colSums
```{r}
dbMatrix::colSums(sparse)

all.equal(unname(dbMatrix::colSums(sparse)), Matrix::colSums(dgc))
```

### rowMeans
```{r}
dbMatrix::rowMeans(sparse)

all.equal(unname(dbMatrix::rowMeans(sparse)), Matrix::rowMeans(dgc))
```

### rowSums
```{r}
dbMatrix::rowSums(sparse)

all.equal(unname(dbMatrix::rowSums(sparse)), Matrix::rowSums(dgc))
```

### dim
```{r}
dim(sparse)

dim(dgc)
```

### Transpose
```{r}
t(sparse)
```

### Dense matrices

We use the same approach as sparse matrices to convert a matrix into a
triplicate `i`, `j`, and `x` vector data.frame format. All values including zeros are included. 

```{r, eval=TRUE, message=FALSE, warning=FALSE}
dense = sim_dbDenseMatrix()

# preview
dense
```
## dbDenseMatrix Operations
### colMeans
```{r}
dbMatrix::colMeans(dense)
```

### colSums
```{r}
dbMatrix::colSums(dense)
```

### rowMeans
```{r}
dbMatrix::rowMeans(dense)
```

### rowSums
```{r}
dbMatrix::rowSums(dense)
```

### mean
```{r}
dbMatrix::mean(dense)
```

### dim
```{r}
dim(dense)
```

### transpose
```{r}
t(dense)
```


## Session Info

```{r eval=TRUE, message=FALSE}
sessionInfo()
```
