---
title: "Quickstart"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
number-sections: true
vignette: >
  %\VignetteIndexEntry{quickstart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

This vignette introduces the different classes of Duckling and how to work 
with them at a basic level using generic functions. The goal of the classes
was to emulate in-memory classes that are commonly used in the omics field
(e.g. dgCMatrix) but rely on a database backend (i.e. DuckDB). 

## Class Structure of Duckling

[![](https://mermaid.ink/img/pako:eNp1VF2PmzAQ_CvIT20vicgngYdKV6WVTuqpUa93Dy1VZMwGrBg7NaYNjfLfu4AJBPWeWM-OZ3fHNmfCVAwkIEzQPN9wmmiahbJeOZuCHQSXiXP-8fLw9dvz_eeflzYXUXYAGT_IvXLOoXScu1jz36B3jAoROCylmjIDuklFuyM16RBOaT7ExhqYkhKY-dAUePM2lNeicbShhjb1ri01FRAPHBOJ3VYp8Zr6nYZMGdhJmsEwxSU3gSNUwnGC21YGPTxSo_mpnZpntVyOe3luWgyXXBpI7FT3mpu0ksH4yzG3EQq1IFPinQ21-tOG5j_Tf8IDAlv8AGWd7xhP6PNWq1NZUWvSwCncBScDEmetuB_ruEmMcwQ4Fb8K0KWt3ClvlSgTJWtxW54aNCIqDHrZ661Oyd0R-T0TLn0lRPO-UMXmVUt9eiiv9288fm8rDMDeLQylvR0d2bZzgzeHV4H2GBsYvdA5dMn-uqVAxoe0G8IGZD_XK3pzKla-Ayypb_DrnKt1ZEQy0BnlMb7e2saQmBRwYBJgGMOeFsKEBL1EKi2MeiolI4HRBYxIccQXA_a9k2BPRY7okcrvSmUtCZckOJMTCaZLf-LOVuvV0pstZvP5ejUiJQlm64nvua43m3srfzH1_PllRP7WCu7En1dsz3eni-XaW4wIxNwo_Wj_N9Xn8g9yVonZ?type=png)](https://mermaid.live/edit#pako:eNp1VF2PmzAQ_CvIT20vicgngYdKV6WVTuqpUa93Dy1VZMwGrBg7NaYNjfLfu4AJBPWeWM-OZ3fHNmfCVAwkIEzQPN9wmmiahbJeOZuCHQSXiXP-8fLw9dvz_eeflzYXUXYAGT_IvXLOoXScu1jz36B3jAoROCylmjIDuklFuyM16RBOaT7ExhqYkhKY-dAUePM2lNeicbShhjb1ri01FRAPHBOJ3VYp8Zr6nYZMGdhJmsEwxSU3gSNUwnGC21YGPTxSo_mpnZpntVyOe3luWgyXXBpI7FT3mpu0ksH4yzG3EQq1IFPinQ21-tOG5j_Tf8IDAlv8AGWd7xhP6PNWq1NZUWvSwCncBScDEmetuB_ruEmMcwQ4Fb8K0KWt3ClvlSgTJWtxW54aNCIqDHrZ661Oyd0R-T0TLn0lRPO-UMXmVUt9eiiv9288fm8rDMDeLQylvR0d2bZzgzeHV4H2GBsYvdA5dMn-uqVAxoe0G8IGZD_XK3pzKla-Ayypb_DrnKt1ZEQy0BnlMb7e2saQmBRwYBJgGMOeFsKEBL1EKi2MeiolI4HRBYxIccQXA_a9k2BPRY7okcrvSmUtCZckOJMTCaZLf-LOVuvV0pstZvP5ejUiJQlm64nvua43m3srfzH1_PllRP7WCu7En1dsz3eni-XaW4wIxNwo_Wj_N9Xn8g9yVonZ)

Duckling currently supports the following generic methods.

|          | dbSparseMatrix | dbDenseMatrix | dbDataFrame |
|----------|----------------|---------------|-------------|
| colSums  | ✅             | ✅            | \-          |
| rowSums  | ✅             | ✅            | \-          |
| colMeans | ✅             | ✅            | \-          |
| rowMeans | ✅             | ✅            | \-          |
| colSds   | 🟧             | ✅            | \-          |
| rowSds   | 🟧             | ✅            | \-          |
| t        | 🟧             | ✅            | \-          |
| mean     | ✅             | ✅            | \-          |
| nrow     | ✅             | ✅            | ✅          |
| ncol     | ✅             | ✅            | ✅          |
| dims     | ✅             | ✅            | ✅          |
| head     | 🟧             | ✅            | \-          |
| tail     | 🟧             | ✅            | \-          |
| math     | 🟧             | ✅            | \-          |
| ...      |                |               |             |

## Loading library

```{r, message=F}
# git clone repo
devtools::load_all('../')
library(dplyr)
library(duckdb)
```

## Working with large matrices with dbMatrix

`dbMatrix` consists of three subclasses that are designed to work with
different types of matrices. All `dbMatrix` subclasses currently support
basic generic operations as listed below, with more to be likely added
in the future.

### Sparse matrices

We represent sparse matrices using a triplicate format with `i`, `j`,
and `x` vectors. `i` and `j` refer to the row and column indices,
respectively, with specific `x` values or counts.

The `dbSparseMatrix` class stores this ijx matrix (now tehnically a
data.frame) with their associated dimension names and dimensions. It
does not contain zeros. Zero values are inferred based on the stored
dimensions.

```{r eval=TRUE, message=FALSE, warning=FALSE}
# simulate dbSparseMatrix
sparse = sim_dbSparseMatrix()

# preview 
dplyr::glimpse(sparse[]) 
```

```{r}
# Generics for dbSparseMatrix
colMeans(sparse)

colSums(sparse)

rowMeans(sparse)

rowSums(sparse)

mean(sparse)

dim(sparse)
```

### Dense matrices

We use the same approach as sparse matrices to convert a matrix into a
triplicate `i`, `j`, and `x` vector data.frame format. All values in the
`x` vector are retained and the `i` and `j` vectors represent the
dimnames.

```{r, eval=TRUE, message=FALSE, warning=FALSE}
dense = sim_dbDenseMatrix() # from MASS::Animals package

dplyr::glimpse(dense[]) 
```

```{r}
colMeans(dense)

colSums(dense)

rowMeans(dense)

rowSums(dense)

mean(dense)

dim(dense)

# t(dense)

# Also supports math operations
# dense + 10
# 
# dense / 10 
# 
# dense * 10
# 
# dense - 10
```

## Working with large data.frames with dbDataFrame

Large data.frames can be stored as dbDataFrames directly in DuckDB since
they are essentially the same tabular data structure.

```{r eval=TRUE, message=FALSE}
# Simulate dbDataFrame
df = sim_dbDataFrame()

# Generics for dbDataFrame
nrow(df)

ncol(df)

dim(df)
```

## Session Info

```{r eval=TRUE, message=FALSE}
sessionInfo()
```
